---
title: "Liver cell segmetnation"
author: "Ludvig Larsson"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(magick)
library(magrittr)
library(EBImage)
library(dplyr)
library(magrittr)
library(ggplot2)
source(file = "../scripts/global_functions.R")

"25800,13500,3600,9700"
im <- image_read(path = "../data/191015_spot_ST_CN65_Liver_D2_0.1.jpg") %>%
  image_crop(geometry = geometry_area(width = 25800, height = 13500, x_off = 3600, y_off = 9700)) %>%
  image_scale(paste0(25800*0.6)) %>%
  as_EBImage()

```

Create a small representative crop window to test segmentation on.

```{r small_crop, fig.width=10, fig.height=10}

im_cropped <- im[5000:6000, 5000:6000, ]
display(im_cropped, method = "raster")

```

There seems to be some blurry areas at the border of the tiles which will be a bit problematic. 

Below is a plot of the three color channels; "red", "green" and "blue". If we focus on the nuclei, it seems as if the red and blue color channels are the most informative. From the histogram we can see that these two channels have a quite similar intensity distribution. We can If we drop the green channel, the nuclei will probably be easier to segment.

```{r channels, fig.width=13, fig.height=3}

par(mfrow = c(1, 4))
hist(im_cropped)
for (col in c("red", "green", "blue")) {
  display(channel(im_cropped, col), method = "raster")
  text(x = 10, y = 20, label = paste0(col, " channel"), adj = c(0, 1), col = "orange", cex = 2)
}

```

Let's remove the green color channel and convert the image to grayscale and then invert the color intensities so that nuclei are bright and the background is dark (just to make things a little easier to work with). I have also added some contrast (gamma factor) to the grayscale image.

```{r grayscale, fig.height=3, fig.width=6.5}

# Create an RGB image with an empty green channel
imrb <- rgbImage(red = channel(im_cropped, "red"), blue = channel(im_cropped, "blue"))
imgray <- channel(imrb, "gray")
imgray <- 1 - imgray

# Add contrast
imgray <- imgray^2

# Plot images
par(mfrow = c(1, 2))
plot(imrb); text(x = 10, y = 20, label = "green channel missing", adj = c(0, 1), col = "orange", cex = 1.5)
plot(imgray); text(x = 10, y = 20, label = "merged RB + inverted in grayscale", adj = c(0, 1), col = "orange", cex = 1.5)

```

## Segmentation workflow

For the segmentation workflow I have created a set of functions that can be used to segment cells based on DAPI or AB staining. However, now that we have nuclei that are pretty clear, it should be possible to use the same principles to segment the nuclei from the H&E stain.

## Thresholding

```{r fig.width = 10, fig.height=10}

nuc <- imgray
#nmask = thresh(nuc, w=5, h=5, offset=0.05) #original setting

f = makeBrush(51, shape='disc', step=FALSE)
f = f/sum(f)
nmask <- nuc > (filter2(nuc, f, boundary="replicate") + 0.08)
#nmask <- nmask > 0.6

nmask = opening(nmask, makeBrush(7, shape='disc')) #original setting
#nmask = opening(nmask, makeBrush(5, shape='disc'))
nmask = fillHull(nmask)

plot(imgray); text(x = 10, y = 20, label = "original image", adj = c(0, 1), col = "orange", cex = 1.5)
plot(nmask); text(x = 10, y = 20, label = "segmented cells", adj = c(0, 1), col = "orange", cex = 1.5)

```

### Label cells

Next we need to label connected "positive" pixels as cells which can be done using `bwlabel`. In the histogram below, we can see the areas of the segmented cells in pixels.

```{r label_cells}
# Next we need to define neighboring sets of "positive" pixels as cells which
# can be done using bwlabel
cells_segmented <- nmask
cells_segmented <- bwlabel(cells_segmented)

hist(table(cells_segmented)[-1], breaks = 100, xlab = "area [pixels]", main = "Cell area [pixels] distribution")

```

### Clean up speckles

```{r clean, fig.width=9.5, fig.height=3}

# Lastly, we remove cells that are either too small or too big
cells_cleaned <- clean_cells(imthreshold = cells_segmented, thr = c(5, 1000))

par(mfrow = c(1, 3))
plot(imgray); text(x = 10, y = 20, label = "original image", adj = c(0, 1), col = "orange", cex = 1.5)
plot(cells_segmented); text(x = 10, y = 20, label = "segmented image", adj = c(0, 1), col = "orange", cex = 1.5)
plot(cells_cleaned); text(x = 10, y = 20, label = "cleaned segmented image", adj = c(0, 1), col = "orange", cex = 1.5)

```

### Watershed

Finally, we can apply the watershed algorithm to try to label nuclei that have been merged


```{r watershed}

cells_watershed <- watershed_cells(imclean = cells_cleaned)
display(colorLabels(cells_watershed), method = "raster")
text(x = 10, y = 20, label = "labelled cells", adj = c(0,1), col = "orange", cex = 1.5)

```

Let's overlay our results on the original image

```{r paint nuclei, fig.width=12, fig.height=12}

cells_colored <- paintObjects(cells_cleaned, im_cropped, col = "#FFA500")
plot(im_cropped); text(x = 10, y = 20, label = "original image", adj = c(0, 1), col = "yellow", cex = 5)
plot(cells_colored); #text(x = 10, y = 20, label = "outlined nuclei", adj = c(0, 1), col = "yellow", cex = 5)

```

## Run segmentation on whole tissue section

```{r segment_whole_section}

seg_cells <- function(im) {
  imrb <- rgbImage(red = channel(im, "red"), blue = channel(im, "blue"))
  imgray <- channel(imrb, "gray")
  imgray <- 1 - imgray
  nuc <- imgray^2
  f = makeBrush(11, shape='disc', step=FALSE)
  f = f/sum(f)
  nmask <- nuc > (filter2(nuc, f, boundary="replicate") + 0.03)
  nmask = opening(nmask, makeBrush(3, shape='disc'))
  nmask = fillHull(nmask)
  cells_segmented <- bwlabel(nmask)
  cells_cleaned <- clean_cells(imthreshold = cells_segmented, thr = c(5, 1000), verbose = TRUE)
  #cells_watershed <- watershed_cells(imclean = cells_cleaned, verbose = TRUE)
  cells_watershed <- cells_cleaned
  return(cells_watershed)
}

# Run workflow
cleaned_cells_section <- seg_cells(im)

```

Export image with painted nuclei

```{r paint_nuclei_export}

cells_colored_section <- paintObjects(cleaned_cells_section, im, col = "#FFA500")
writeImage(x = cells_colored_section, type = "jpeg", files = "../images/outlined_cells.jpeg", quality = 100)

```

```{r colorlabels}

writeImage(x = colorLabels(cleaned_cells_section), type = "jpeg", files = "../images/labelled_cells.jpeg", quality = 100)

```

Export separate layers

```{r export_HE}

writeImage(x = im, type = "jpeg", files = "../images/HE_cropped.jpeg", quality = 100)
emptyim <- Image(data = array(0, dim = c(8000, 8678)), colormode = Color)
cells_colored_section <- paintObjects(cleaned_cells_section, emptyim, col = "#FFA500")
writeImage(x = cells_colored_section, type = "jpeg", files = "../images/labelled_cells_only.jpeg", quality = 100)

```

## Load spot selection files

```{r spots}

spot_coordinates <- read.table(file = "../data/C1_cropped_180rot.tsv", header = T, sep = "\t")

```

## Extract cell features

```{r cell_features, fig.width=8, fig.height=8}

# Extract centroid coordinates for cells
fts.moment <- computeFeatures.moment(cleaned_cells_section)

# Create a cell coordinate table 
cell_coordinates <- data.frame(fts.moment[, c("m.cx", "m.cy")])

# To make the coordinates compatible with the spot selection table,
# we need to adjust them to account for the rescaling and cropping 
# done in the first step
original_HE_width <- 10325 # width of HE used for ST spot detector (pixels)
downscaled_HE_width <- 8000 # width of downscaled HE image used for this segmentation process
sf <- original_HE_width/downscaled_HE_width # Scalefactor

# Convert cell coordinates to fit to the same coordinate system as the spots
cell_coordinates$m.cx <- ((cell_coordinates$m.cx*sf) + 2250)*1#0.1
cell_coordinates$m.cy <- ((cell_coordinates$m.cy*sf) + 1015)*1#0.1

ggplot() +
  geom_point(data = cell_coordinates, aes(m.cx, m.cy), size = 0.1) +
  geom_point(data = spot_coordinates, aes(pixel_x, pixel_y), color = "red") +
  theme_void() +
  scale_x_continuous(limits = c(0, 30131/2)) +
  scale_y_continuous(limits = c(0, 27755/2))

```

## Compute minimum spot-spot distance

```{r spot_dist}

distMat <- dist(spot_coordinates[, c("pixel_x", "pixel_y")])
distMat <- as.matrix(distMat)
diag(distMat) <- Inf
center_to_center_distance <- mean(apply(distMat, 2, min))

```

## Get cells within spots

Label cells that are within 1 spot radius of the spot centers.

```{r cells_within_spots}

spot_radius <- center_to_center_distance*0.25

# Calculate pairwise distances
set1 <- spot_coordinates[, c("pixel_x", "pixel_y")]
set2 <- cell_coordinates
mindist <- apply(set1, 1, function(x) {
  sqrt(colSums((t(set2[, 1:2]) - x)^2))
})
spotids <- paste0(spot_coordinates$x, "x", spot_coordinates$y)

arrind <- which(mindist <= spot_radius, arr.ind = TRUE)
df <- data.frame(dist = apply(arrind, 1, function(x) {mindist[x[1], x[2]]}),
           ID = arrind[, "row"], spot = spotids[arrind[, "col"]])
cell_coordinates$within <- "no"
cell_coordinates$within[df$ID] <- "yes"
cell_coordinates$spot <- NA
cell_coordinates$spot[df$ID] <- df$spot

```

Plot selected cells

```{r plot_cells, fig.width=8, fig.height=8}

ggplot(cell_coordinates, aes(m.cx, 3174 - m.cy, color = within)) +
  geom_point(size = 0.1) +
  theme_void() +
  scale_x_continuous(limits = c(0, 3174), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 3174), expand = c(0, 0))

```
Count number of cells per spot

```{r count_cells_per_spot}

cell_coordinates_subset <- na.omit(cell_coordinates)
cells_per_spot <- cell_coordinates_subset %>% 
  group_by(within, spot) %>%
  summarize(count = n())

```


Create a new cell segmentation object where cells outside of spots are removed


```{r remove_cells}

all_inds <- 1:98084
rm_inds <- setdiff(all_inds, df$ID)
cleaned_cells_section_filtered <- rmObjects(cleaned_cells_section, index = rm_inds, reenumerate = FALSE)
writeImage(x = cleaned_cells_section_filtered, type = "jpeg", files = "../images/cells_on_spots.jpeg", quality = 100)

```

## Date

```{r date}
date
```

## Session info

```{r session}
sessionInfo()
```
